Need Fulfillment During Intergroup Contact: Three Experience Sampling Studies

### 2023-05-31 Methods-and-Results.Rmd
robustness re-render with descr
interaction qual cor added

\subsubsection{Participants} After receiving ethical approval from the University Masked for Peer Review, we recruited `r nrow(workerSampleInfo)` non-Dutch migrants using the local paid participant pool. Participants reported on their interactions for at least 30 days with two daily measures (capturing the morning and afternoon). With this design, we aimed at getting 50-60 measurements per participant (\textit{M} = `r format(round(mean(workerSampleInfo$dailiesN),2), nsmall=2)`, \textit{SD} = `r format(round(sd(workerSampleInfo$dailiesN),2), nsmall=2)`, \textit{total N} = `r format(sum(workerSampleInfo$dailiesN), big.mark=",")`). This is a common number of measurements found in experience sampling studies and offers sufficient power to model processes within and between participants \citep[e.g.,][]{AanhetRot2012}. Participants were compensated for their participation with up to 34 Euros – each two Euros for pre- and post-questionnaire and 50 Eurocents for every experience sampling measurement. The sample consisted of relatively young, educated, and western migrants from the global north ($M_{age}$ = `r format(round(mean(workerSampleInfo$age),2), nsmall=2)`, $SD_{age}$ = `r format(round(sd(workerSampleInfo$age),2), nsmall=2)`, `r nrow(workerSampleInfo %>% filter(gender == "Female"))` women, `r nrow(workerSampleInfo %>% filter(occupation == "student"))` students). The sample accurately describes the largest groups of migrants in the region \citep[][]{GemeenteGroningen2015}.

@@

\paragraph{Perceived Interaction Quality}To assess ratings of the perceived interaction quality, participants rated the statement “\textit{Overall the interaction was …}” on two continuous slider scales measuring pleasantness \citep[from unpleasant (-50) to pleasant (+50)) and meaningfulness (from superficial (-50) to meaningful (+50); both items adapted from][]{Downie2008}.

@@

\subsubsection{Contact Hypothesis}
Using a multilevel regression, we find that having an outgroup contact is indeed associated with significantly more positive outgroup attitudes (`r MlCoeffLatex(lmeMdl = mdlWorker$lmeInterceptAttType, lmerCI = mdlWorker$lmerInterceptAttTypeCI, varName = "OutgroupInteractionC")`), even after controlling for having an interaction with a non-Dutch (which did not relate to outgroup attitudes independently). Additionally, while multi-level regressions are generally robust against unequal cell sizes, we correct for inequalities by using centered predictors and reintroducing the means as level two predictors (\citealp{yaremych2021a}; for full results see \tblref{tab:intergroupGeneralTblLong}, \fgrref{fig:ContactHypothesis}, and Online Supplementary Material A)\footnote{Interestingly, adding random slopes to this model did not explain additional variance. This is unusual and might indicate that the effect is very consistent across participants. However, the small number of participants, or other measurement issues provide an alternative explanation, which is why we offer a combined data set analysis as part of our stability analyses.}. 
Thus, in our first data, we find initial evidence that outgroup contacts show a positive effect on outgroup attitudes within real-life data.

@@

\subsubsection{Participants} We recruited `r nrow(studentSampleInfo)` international students using a local participant pool. We specifically targeted non-Dutch students, who had recently arrived in the Netherlands. Participants reported on their interactions for at least 30 days with two daily measures (capturing the morning and afternoon). With this design, we again aimed at receiving 50-60 measurements per participant (\textit{M} = `r format(round(mean(studentSampleInfo$dailiesN),2), nsmall=2)`, \textit{SD} = `r format(round(sd(studentSampleInfo$dailiesN),2), nsmall=2)`, \textit{total N} = `r format(sum(studentSampleInfo$dailiesN), big.mark=",")`). As with the previous study, this should offer sufficient power to model processes within participants and will lend stronger weight to between-participant results. Participants were compensated for their participation with partial course credits --- depending on their participation. The sample consisted of relatively young migrants, who were mostly from the global north ($M_{age}$ = `r format(round(mean(studentSampleInfo$age),2), nsmall=2)`, $SD_{age}$ = `r format(round(sd(studentSampleInfo$age),2), nsmall=2)`, `r nrow(studentSampleInfo %>% filter(gender == "women"))` women). The sample fairly accurately describes the local population of international students.

@@

\paragraph{Perceived Interaction Quality}The ratings of the perceived contact quality were identical to Study 1.

@@

\subsubsection{Participants} We recruited `r nrow(medicalSampleInfo)` international medical students using contacts within the University Medical School. We specifically targeted non-Dutch students, who had recently arrived in the Netherlands. Participants reported on their interactions for at least 30 days with two daily measures (capturing the morning and afternoon). With this design, we aimed at getting 50-60 measurements per participant (\textit{M} = `r format(round(mean(medicalSampleInfo$dailiesN),2), nsmall=2)`, \textit{SD} = `r format(round(sd(medicalSampleInfo$dailiesN),2), nsmall=2)`, \textit{total N} = `r format(sum(medicalSampleInfo$dailiesN), big.mark=",")`). As with the previous studies, this offered sufficient power to model processes within participants. Participants were compensated in the same manner as during Study 1. The sample consisted of relatively young migrants ($M_{age}$ = `r format(round(mean(medicalSampleInfo$age),2), nsmall=2)`, $SD_{age}$ = `r format(round(sd(medicalSampleInfo$age),2), nsmall=2)`, `r nrow(medicalSampleInfo %>% filter(gender == "women"))` women). The sample fairly accurately describes the local population of young international medical professionals.

@@

\paragraph{Psychological Needs}The measurement of the core situational need and its fulfillment was identical to Study 2.

@@

@@

\subsubsection{Interaction intent}
To test whether the need mechanism was affected by whether the interaction was accidental or planned we ran an exploratory moderation analysis using the participants' ratings of how much they perceived the interaction as 'accidental'. It should be noted that, in order to keep the ESM surveys short, we asked our participants to focus on most the important interaction (i.e., "\textit{The following questions will be about the interaction \underline{you consider most significant}.}"; emphasis as in original). This was also reflected in a relatively low mean and consistent right skew of the 'accidental' item. Nonetheless, there remained a substantial variance in the item and we continued with the moderation analysis. We found that both in the overall test as in all studies individually key need fulfillment remained a strong predictor of outgroup attitudes, even when accounting for differences in whether the interaction was accidental (rather than planned). Moreover, in none of the analysis did we see a moderation effect of interaction intent nor did we find a main effect of interaction intent. There is, thus, consistent evidence that the need fulfillment mechanism was not meaningfully different for more accidental interactions.

\subsubsection{Well-being outcome}
Given the well-established criticism that more positive outgroup attitudes might not always be beneficial for minority group members \citep[e.g.,][]{reimer2023}, we conducted an additional exploratory analysis assessing the effect of need fulfilling outgroup interactions on reported well-being. Experienced well-being is a common indicator of health and an important life quality measurement in itself, especially for migrants or other minority groups \citep[e.g.,][]{bhugra2011}. We find that the results with well-being as the outcome variable mirror our main results in effect size and interpretation. The results are consistent across and within studies and, thus, add weight to the importance of considering the need fulfillment experiences when it comes to outgroup interactions.

>>>>

\subsubsection{Participants} After receiving ethical approval from the University Masked for Peer Review, we recruited `r nrow(workerSampleInfo)` non-Dutch migrants using the local paid participant pool. Participants reported on their interactions for at least 30 days with two daily measures (capturing the morning and afternoon). With this design, we aimed at getting 50-60 measurements per participant (\textit{M} = `r format(round(mean(workerSampleInfo$dailiesN),2), nsmall=2)`, \textit{SD} = `r format(round(sd(workerSampleInfo$dailiesN),2), nsmall=2)`, \textit{total N} = `r format(sum(workerSampleInfo$dailiesN), big.mark=",")`). This is a common number of measurements found in experience sampling studies and offers sufficient power to model processes within and between participants \citep[e.g.,][]{AanhetRot2012}. Participants were compensated for their participation with up to 34 Euros – each two Euros for pre- and post-questionnaire and 50 Eurocents for every experience sampling measurement. The sample consisted of relatively young, educated, and western migrants from the global north ($M_{age}$ = `r format(round(mean(workerSampleInfo$age),2), nsmall=2)`, $SD_{age}$ = `r format(round(sd(workerSampleInfo$age),2), nsmall=2)`, `r nrow(workerSampleInfo %>% filter(gender == "Female"))` women, `r nrow(workerSampleInfo %>% filter(occupation == "student"))` students). The sample accurately describes the largest groups of migrants in the region \citep[see][for a recent report on the largest migrant groups and see Supplemental Material A for a full overview of the demographic composition, including country of origin]{GemeenteGroningen2015}.

@@

```{r}
#| label: s1_quality_cor
#| include: false
library(correlation)

s1_quality_cor <- misty::multilevel.cor(
  dtWorkerSupp$workerInteractionType %>%
    filter(OutgroupInteractionLab == "Yes") %>%
    select(quality_overall_1, quality_meaning_1),
  dtWorkerSupp$workerInteractionType %>%
    filter(OutgroupInteractionLab == "Yes") %>%
    select(PID),
  output = FALSE
)
```

\paragraph{Perceived Interaction Quality}To assess ratings of the perceived interaction quality, participants rated the statement “\textit{Overall the interaction was …}” on two continuous slider scales measuring pleasantness (from unpleasant (-50) to pleasant (+50)) and meaningfulness (from superficial (-50) to meaningful (+50)). The items formed a coherent concept within the participants ($r_{within}$ = `r s1_quality_cor$result$with.cor[2] %>% round(2) %>% format(., nsmall = 2)`, \textit{p} `r ifelse(s1_quality_cor$result$with.p[3] < 0.001, "< .001", paste0("= ", s1_quality_cor$result$with.p[3] %>% round(3) %>% format(., nsmall = 3)))`). We adapted the from \citet[][]{Downie2008}, who validated the approach.

@@

\subsubsection{Contact Hypothesis}
Using a multilevel regression, we find that having an outgroup contact is indeed associated with significantly more positive outgroup attitudes (`r MlCoeffLatex(lmeMdl = mdlWorker$lmeInterceptAttType, lmerCI = mdlWorker$lmerInterceptAttTypeCI, varName = "OutgroupInteractionC")`), even after controlling for having an interaction with a non-Dutch (which did not relate to outgroup attitudes independently). Additionally, while multilevel regressions are generally robust against unequal cell sizes, we correct for inequalities by using centered predictors and reintroducing the means as level two predictors (\citealp{yaremych2021a}; for full results see \tblref{tab:intergroupGeneralTblLong}, \fgrref{fig:ContactHypothesis}, and Online Supplementary Material A)\footnote{Interestingly, adding random slopes to this model did not explain additional variance. This is unusual and might indicate that the effect is very consistent across participants. However, the small number of participants, or other measurement issues provide an alternative explanation, which is why we offer a combined data set analysis as part of our stability analyses.}. 
Thus, in our first data, we find initial evidence that outgroup contacts show a positive effect on outgroup attitudes within real-life data.

@@

\subsubsection{Participants} We recruited `r nrow(studentSampleInfo)` international students using a local participant pool. We specifically targeted non-Dutch students, who had recently arrived in the Netherlands. Participants reported on their interactions for at least 30 days with two daily measures (capturing the morning and afternoon). With this design, we again aimed at receiving 50-60 measurements per participant (\textit{M} = `r format(round(mean(studentSampleInfo$dailiesN),2), nsmall=2)`, \textit{SD} = `r format(round(sd(studentSampleInfo$dailiesN),2), nsmall=2)`, \textit{total N} = `r format(sum(studentSampleInfo$dailiesN), big.mark=",")`). As with the previous study, this should offer sufficient power to model processes within participants and will lend stronger weight to between-participant results. Participants were compensated for their participation with partial course credits --- depending on their participation. The sample consisted of relatively young migrants, who were mostly from the global north ($M_{age}$ = `r format(round(mean(studentSampleInfo$age),2), nsmall=2)`, $SD_{age}$ = `r format(round(sd(studentSampleInfo$age),2), nsmall=2)`, `r nrow(studentSampleInfo %>% filter(gender == "women"))` women). The sample fairly accurately describes the local population of international students (see Supplemental Materials A for additional demographic information).

@@

```{r}
#| label: s2_quality_cor
#| include: false

s2_quality_cor <- misty::multilevel.cor(
  dtStudentSupp$studentInteractionType %>%
    filter(OutgroupInteractionLab == "Yes") %>%
    select(quality_overall, quality_meaning),
  dtStudentSupp$studentInteractionType %>%
    filter(OutgroupInteractionLab == "Yes") %>%
    select(PID),
  output = FALSE
)
```

\paragraph{Perceived Interaction Quality}The ratings of the perceived contact quality were identical to Study 1 (item correlation: $r_{within}$ = `r s2_quality_cor$result$with.cor[2] %>% round(2) %>% format(., nsmall = 2)`, \textit{p} `r ifelse(s2_quality_cor$result$with.p[3] < 0.001, "< .001", paste0("= ", s2_quality_cor$result$with.p[3] %>% round(3) %>% format(., nsmall = 3)))`).

@@

\subsubsection{Participants} We recruited `r nrow(medicalSampleInfo)` international medical students using contacts within the University Medical School. We specifically targeted non-Dutch students, who had recently arrived in the Netherlands. Participants reported on their interactions for at least 30 days with two daily measures (capturing the morning and afternoon). With this design, we aimed at getting 50-60 measurements per participant (\textit{M} = `r format(round(mean(medicalSampleInfo$dailiesN),2), nsmall=2)`, \textit{SD} = `r format(round(sd(medicalSampleInfo$dailiesN),2), nsmall=2)`, \textit{total N} = `r format(sum(medicalSampleInfo$dailiesN), big.mark=",")`). As with the previous studies, this offered sufficient power to model processes within participants. Participants were compensated in the same manner as during Study 1. The sample consisted of relatively young migrants ($M_{age}$ = `r format(round(mean(medicalSampleInfo$age),2), nsmall=2)`, $SD_{age}$ = `r format(round(sd(medicalSampleInfo$age),2), nsmall=2)`, `r nrow(medicalSampleInfo %>% filter(gender == "women"))` women). The sample fairly accurately describes the local population of young international medical professionals (see Supplemental Materials A for additional demographic information).

@@

\paragraph{Need Fulfillment}The measurement of the core situational need and its fulfillment was identical to Study 2.

@@

```{r}
#| label: s3_quality_cor
#| include: false

s3_quality_cor <- misty::multilevel.cor(
  dtMedical$full %>%
    filter(OutgroupInteractionLab == "Yes") %>%
    select(qualityOverall, qualityMeaning),
  dtMedical$full %>%
    filter(OutgroupInteractionLab == "Yes") %>%
    select(PID),
  output = FALSE
)
```

\paragraph{Perceived Interaction Quality}The ratings of the perceived contact quality were identical to Study 1 and 2 (item correlation: $r_{within}$ = `r s3_quality_cor$result$with.cor[2] %>% round(2) %>% format(., nsmall = 2)`, \textit{p} `r ifelse(s3_quality_cor$result$with.p[3] < 0.001, "< .001", paste0("= ", s3_quality_cor$result$with.p[3] %>% round(3) %>% format(., nsmall = 3)))`).

@@

\subsubsection{Interaction intent}
To test whether the need mechanism was affected by whether the interaction was accidental or planned we ran an exploratory moderation analysis using the participants' ratings of how much they perceived the interaction as 'accidental'. It should be noted that, in order to keep the ESM surveys short, we asked our participants to focus on most the important interaction (i.e., "\textit{The following questions will be about the interaction \underline{you consider most significant}.}"; emphasis as in original). This was also reflected in a relatively low mean and consistent right skew of the 'accidental' item. Nonetheless, there remained a substantial variance in the item and we continued with the moderation analysis. We found that both in the overall test and in all studies individually key need fulfillment remained a strong predictor of outgroup attitudes, even when accounting for differences in whether the interaction was accidental (rather than planned). Moreover, in none of the analyses did we see a moderation effect of interaction intent nor did we find a main effect of interaction intent. There is, thus, consistent evidence that the need fulfillment mechanism was not meaningfully different for more accidental interactions.

\subsubsection{Well-being outcome}
Given the well-established criticism that more positive outgroup attitudes might not always be beneficial for minority group members \citep[e.g.,][]{reimer2023}, we conducted an additional exploratory analysis assessing the effect of need-fulfilling outgroup interactions on reported well-being. Experienced well-being is a common indicator of health and an important life quality measurement in itself, especially for migrants or other minority groups \citep[e.g.,][]{bhugra2011}. We find that the results with well-being as the outcome variable mirror our main results in effect size and interpretation. The results are consistent across and within studies and, thus, add weight to the importance of considering the need fulfillment experiences when it comes to outgroup interactions.

### 2023-05-31 Methods-and-Results.tex
interaction qual cor added

To assess ratings of the perceived interaction quality, participants
rated the statement ``\textit{Overall the interaction was …}'' on two
continuous slider scales measuring pleasantness
\citep[from unpleasant (-50) to pleasant (+50)) and meaningfulness (from superficial (-50) to meaningful (+50); both items adapted from][]{Downie2008}.

@@

We recruited 113 international students using a local participant pool.
We specifically targeted non-Dutch students, who had recently arrived in
the Netherlands. Participants reported on their interactions for at
least 30 days with two daily measures (capturing the morning and
afternoon). With this design, we again aimed at receiving 50-60
measurements per participant (\textit{M} = 43.94, \textit{SD} = 15.00,
\textit{total N} = 4,965). As with the previous study, this should offer
sufficient power to model processes within participants and will lend
stronger weight to between-participant results. Participants were
compensated for their participation with partial course credits ---
depending on their participation. The sample consisted of relatively
young migrants, who were mostly from the global north (\(M_{age}\) =
20.24, \(SD_{age}\) = 2.12, 84 women). The sample fairly accurately
describes the local population of international students.

@@

The ratings of the perceived contact quality were identical to Study 1.

@@

\paragraph{Psychological Needs}

@@

>>>>

To assess ratings of the perceived interaction quality, participants
rated the statement ``\textit{Overall the interaction was …}'' on two
continuous slider scales measuring pleasantness (from unpleasant (-50)
to pleasant (+50)) and meaningfulness (from superficial (-50) to
meaningful (+50)). The items formed a coherent concept within the
participants (\(r_{within}\) = 0.54, \textit{p} \textless{} .001). We
adapted the from \citet[][]{Downie2008}, who validated the approach.

@@

We recruited 113 international students using a local participant pool.
We specifically targeted non-Dutch students, who had recently arrived in
the Netherlands. Participants reported on their interactions for at
least 30 days with two daily measures (capturing the morning and
afternoon). With this design, we again aimed at receiving 50-60
measurements per participant (\textit{M} = 43.94, \textit{SD} = 15.00,
\textit{total N} = 4,965). As with the previous study, this should offer
sufficient power to model processes within participants and will lend
stronger weight to between-participant results. Participants were
compensated for their participation with partial course credits ---
depending on their participation. The sample consisted of relatively
young migrants, who were mostly from the global north (\(M_{age}\) =
20.24, \(SD_{age}\) = 2.12, 84 women). The sample fairly accurately
describes the local population of international students (see
Supplemental Materials A for additional demographic information).

@@

The ratings of the perceived contact quality were identical to Study 1
(item correlation: \(r_{within}\) = 0.39, \textit{p} \textless{} .001).

@@

@@

The ratings of the perceived contact quality were identical to Study 1
and 2 (item correlation: \(r_{within}\) = 0.46, \textit{p} \textless{}
.001).
